Next Generation Planning - Structuring and Sharing Environmental Drivers Data for the St. Lawrence System

### 2020-03-11 Revisions/Beauchesne-eDriversMS.Rmd
update references

\textcolor{blue}{Holistic approaches typically involve, but are not limited to, the following steps: selecting and describing valued ecosystem components (\textit{e.g.} habitats and species) and drivers (\textit{e.g.} marine traffic and ocean acidification), assessing the exposure and vulnerability of valued components to drivers, selecting a proper spatio-temporal scale, monitoring, and public and stakeholder participation (Dubé and Munkittrick 2001).} <!-- [@dube2001] jones2016 ???? Noble2013??? -->
Gathering environmental
\textcolor{red}{\sout{data for large scale, systematic}}
\textcolor{blue}{knowledge for holistic}
initiatives can, however, be a very challenging and time consuming – not to say painful – process. On one hand, there is an overwhelming and expanding wealth of data available. Such information overload may inhibit our ability to make decisions based on scientific information, promote massive effort duplication, disproportionately appropriate research funds to certain sectors, and obscure knowledge gaps amid a sea of information [@eppler2004]. On the other hand, crucial data are lacking and remain largely unavailable or inaccessible for a variety of reasons, including proprietary rights, lack of organizational time, capacity and training, and in some \textcolor{red}{\sout{rare}} cases unwillingness to share, curtailing our ability for appropriate decision-making.

@@

\textcolor{red}{\sout{While the focus of our paper has been on the description of drivers in the St. Lawrence, drivers are but one of the 4 elements required to properly evaluate environmental impacts. This knowledge node should be weaved with other, comparable, nodes required for integrated management approaches such as species distribution and marine protected areas. Ultimately, all of these knowledges could be weaved together through social-ecological metanetworks (Dee et al. 2017).}}
\textcolor{blue}{The next step will be the inclusion of other types of knowledge to our initiative. Our focus has been on a single element required for fully operational impact assessments. Data that provide knowledge on the exposure of ecosystems to drivers are called stressor-based indicators (Dube and Munkittrick 2001; Dubé 2003; Jones 2016). These indicators efficently identify potential local impacts and can be proactively linked to decision-making, yet assume complete knowledge of drivers but fail to diagnose impacts on valued components or non-additive effects (Dube and Munkittrick 2001; Dubé 2003). In contrast, effect-based indicators are direct measurements of valued components (\textit{e.g.} species abundance and biodiversity) and inherently capture the effects of multiple drivers (Dube and Munkittrick 2001; Dubé 2003). Whereas effect-based indicators are considered superior to stressor-based indicators, they fail to ascribe observed effects to specific drivers (Dube and Munkittrick 2001; Dubé 2003). Stressor-based and effect-based indicators are, therefore, both required to diagnose causes of ecosystem change (Dube 2003; Jones 2016). As a collection of knowledge on stressor-based indicators, \textit{eDrivers} should be weaved with other, comparable, collections of knowledge describing valued ecosystem components that can be linked to drivers and allow for a better understanding of cumulative impacts. Ultimately, interdisciplinary colections of knowledge could be weaved together through social-ecological metanetworks analyses (Dee et al. 2016).}

@@

<!-- !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!! -->
TODO: Remove [@dube2001; @dube2003; @jones2016]
<!-- !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!! -->

>>>>

\textcolor{blue}{Holistic approaches typically involve, but are not limited to, the following steps: selecting and describing valued ecosystem components (\textit{e.g.} habitats and species) and drivers (\textit{e.g.} marine traffic and ocean acidification), assessing the exposure and vulnerability of valued components to drivers, selecting a proper spatio-temporal scale, monitoring, and public and stakeholder participation} [@dube2001].
Gathering environmental
\textcolor{red}{\sout{data for large scale, systematic}}
\textcolor{blue}{knowledge for holistic}
initiatives can, however, be a very challenging and time consuming – not to say painful – process. On one hand, there is an overwhelming and expanding wealth of data available. Such information overload may inhibit our ability to make decisions based on scientific information, promote massive effort duplication, disproportionately appropriate research funds to certain sectors, and obscure knowledge gaps amid a sea of information [@eppler2004]. On the other hand, crucial data are lacking and remain largely unavailable or inaccessible for a variety of reasons, including proprietary rights, lack of organizational time, capacity and training, and in some \textcolor{red}{\sout{rare}} cases unwillingness to share, curtailing our ability for appropriate decision-making.

@@

\textcolor{red}{\sout{While the focus of our paper has been on the description of drivers in the St. Lawrence, drivers are but one of the 4 elements required to properly evaluate environmental impacts. This knowledge node should be weaved with other, comparable, nodes required for integrated management approaches such as species distribution and marine protected areas. Ultimately, all of these knowledges could be weaved together through social-ecological metanetworks (Dee et al. 2017).}}
\textcolor{blue}{The next step will be the inclusion of other types of knowledge to our initiative. Our focus has been on a single element required for fully operational impact assessments. Data that provide knowledge on the exposure of ecosystems to drivers are called stressor-based indicators} [@dube2001; @dube2003; @jones2016]. \textcolor{blue}{These indicators efficently identify potential local impacts and can be proactively linked to decision-making, yet assume complete knowledge of drivers but fail to diagnose impacts on valued components or non-additive effects} [@dube2001; @dube2003]. \textcolor{blue}{In contrast, effect-based indicators are direct measurements of valued components (\textit{e.g.} species abundance and biodiversity) and inherently capture the effects of multiple drivers} [@dube2001; @dube2003]. \textcolor{blue}{Whereas effect-based indicators are considered superior to stressor-based indicators, they fail to ascribe observed effects to specific drivers} [@dube2001; @dube2003]. \textcolor{blue}{Stressor-based and effect-based indicators are, therefore, both required to diagnose causes of ecosystem change} [@dube2003; @jones2016]. \textcolor{blue}{As a collection of knowledge on stressor-based indicators, \textit{eDrivers} should be weaved with other, comparable, collections of knowledge describing valued ecosystem components that can be linked to drivers and allow for a better understanding of cumulative impacts. Ultimately, interdisciplinary colections of knowledge could be weaved together through social-ecological metanetworks analyses} [@dee2017].

@@
